---
title: "Top Hits Spotify"
author: "Yixin Wang"
output: 
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## An Overview of Dataset

The dataset is downloaded as a csv file and is obtained from (https://www.kaggle.com/datasets/paradisejoy/top-hits-spotify-from-20002019?resource=download&select=songs_normalize.csv)

```{r}
top_hits <- read.csv("songs_normalize.csv")
summary(top_hits)
```

In summary, there are 2000 observations and 18 predictors. I think I will work with numbers mostly and there is no missing data.

## An Overview of Research Questions

In this dataset, I am interested in predicting the popularity of these songs. The popularity is between 0 and 89. I think the energy, danceability, loudness, speechiness, liveness, acoustcness would all be useful in predicting popularity and they should be positively related. 

## Proposed Project Timeline

Since my data memo finished late, I am going to start the exploratory data analysis after the data memo passed immediately.

## Questions or Concerns

I think everything is fine for now. 