---
title: "Homework 5"
author: "Yixin Wang"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Elastic Net Tuning

The goal of this assignment is to build a statistical learning model that can predict the **primary type** of a Pokémon based on its generation, legendary status, and six battle statistics.

Read in the file and familiarize yourself with the variables using `pokemon_codebook.txt`.

```{r}
library(tidymodels)
library(tidyverse)
library(discrim)
library(poissonreg)
library(ISLR)
library(ISLR2)
library(corrr)
library(klaR)
library(glmnet)
tidymodels_prefer()
set.seed(3435)
```

### Exercise 1

Install and load the `janitor` package. Use its `clean_names()` function on the Pokémon data, and save the results to work with for the rest of the assignment. What happened to the data? Why do you think `clean_names()` is useful?

```{r}
library(janitor)
```

```{r}
pokemon <- read_csv(file = "Pokemon.csv")
pokemon <- clean_names(pokemon)
pokemon <- as_tibble(pokemon)
pokemon
```

The data handles problematic variable names. The clean_names() function cleans the variable names and return to data frame.

### Exercise 2

Using the entire data set, create a bar chart of the outcome variable, `type_1`.

```{r}
library(ggplot2)
ggplot(pokemon, aes(x = type_1)) + geom_bar()
```

How many classes of the outcome are there? Are there any Pokémon types with very few Pokémon? If so, which ones?

There are 18 classes of the outcome. The flying and fairy types are have few Pokemons.

For this assignment, we'll handle the rarer classes by simply filtering them out. Filter the entire data set to contain only Pokémon whose `type_1` is Bug, Fire, Grass, Normal, Water, or Psychic.

After filtering, convert `type_1` and `legendary` to factors.

```{r}
pokemon <- pokemon[pokemon$type_1 %in% c("Bug","Fire","Grass","Normal","Water","Psychic"), ]
```

```{r}
pokemon$type_1 <- as.factor(pokemon$type_1)
pokemon$legendary <- as.factor(pokemon$legendary)
```

### Exercise 3

Perform an initial split of the data. Stratify by the outcome variable. You can choose a proportion to use. Verify that your training and test sets have the desired number of observations.

Next, use *v*-fold cross-validation on the training set. Use 5 folds. Stratify the folds by `type_1` as well. *Hint: Look for a `strata` argument.* Why might stratifying the folds be useful?

```{r}
pokemon_split <- initial_split(pokemon, prop = 0.7, strata = type_1)
pokemon_train <- training(pokemon_split)
pokemon_test <- testing(pokemon_split)
```

```{r}
dim(pokemon_train)
```

```{r}
dim(pokemon_test)
```

```{r}
pokemon_fold <- vfold_cv(pokemon_train, v = 5, strata = type_1)
```

Since the outcome variable is imbalanced, there are less observations for flying and fairy than other types.

### Exercise 4

Set up a recipe to predict `type_1` with `legendary`, `generation`, `sp_atk`, `attack`, `speed`, `defense`, `hp`, and `sp_def`.

- Dummy-code `legendary` and `generation`;

- Center and scale all predictors.

```{r}
pokemon_recipe <- recipe(type_1 ~ legendary + generation + sp_atk + attack + speed + defense + hp + sp_def, pokemon_train) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_scale(all_predictors()) %>%
  step_center(all_predictors())
```

### Exercise 5

We'll be fitting and tuning an elastic net, tuning `penalty` and `mixture` (use `multinom_reg` with the `glmnet` engine).

Set up this model and workflow. Create a regular grid for `penalty` and `mixture` with 10 levels each; `mixture` should range from 0 to 1. For this assignment, we'll let `penalty` range from -5 to 5 (it's log-scaled).

```{r}
pokemon_spec <- multinom_reg(penalty = tune(), mixture = tune()) %>%
  set_engine("glmnet") %>%
  set_mode("classification")

pokemon_grid <- grid_regular(penalty(c(-5, 5)), mixture(c(0,1)), levels = 10)

pokemon_workflow <- workflow() %>%
  add_recipe(pokemon_recipe) %>%
  add_model(pokemon_spec)
```

How many total models will you be fitting when you fit these models to your folded data?

There are 10 models will be fitted to 5 folded data.

### Exercise 6

Fit the models to your folded data using `tune_grid()`.

```{r}
tune_res <- tune_grid(pokemon_workflow, resamples = pokemon_fold, grid = pokemon_grid)
```

Use `autoplot()` on the results. What do you notice? Do larger or smaller values of `penalty` and `mixture` produce better accuracy and ROC AUC?

```{r}
autoplot(tune_res)
```

The smaller the penalty lead to the better accuracy and ROC AUC. Similarly, the smaller mixture leads to the better accuracy and ROC AUC.

### Exercise 7

Use `select_best()` to choose the model that has the optimal `roc_auc`. Then use `finalize_workflow()`, `fit()`, and `augment()` to fit the model to the training set and evaluate its performance on the testing set.

```{r}
pokemon_penalty <- select_best(tune_res, metric = "roc_auc")
```

```{r}
final = pokemon_workflow %>%
  finalize_workflow(pokemon_penalty) %>%
  fit(pokemon_train) %>%
  augment(pokemon_test)

accuracy(final, truth = type_1, estimate = .pred_class)
```

### Exercise 8

Calculate the overall ROC AUC on the testing set.

```{r}
roc_auc(final, truth = type_1, .pred_Bug, .pred_Fire, .pred_Grass, .pred_Normal, .pred_Water, .pred_Psychic)
```

Then create plots of the different ROC curves, one per level of the outcome. Also make a heat map of the confusion matrix.

What do you notice? How did your model do? Which Pokemon types is the model best at predicting, and which is it worst at? Do you have any ideas why this might be?

```{r}
autoplot(roc_curve(final, truth = type_1, .pred_Bug, .pred_Fire, .pred_Grass, .pred_Normal, .pred_Water, .pred_Psychic))
```

```{r}
conf_mat(final, truth = type_1, estimate = .pred_class) %>%
  autoplot(type = "heatmap") + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The roc_auc is approximately 0.6072. The accuracy is approximately 0.45. I notice that the model is not good at predicting the type of Pokemon. However, the normal type is the model best at predicting. The psychic type is the worst type to predict in the graph. I think the reason is probably that the psychic is quite different than the other Pokemons. 
